---
title: "Untitled"
output: html_document
date: "2024-05-16"
---
# install and loading packages
```{r}
install.packages("sf") 
install.packages("terra")
install.packages("tmap")
```
#load into R library
```{r}
library(tidyverse)
library(sf) # simple features
library (terra) # for raster
library(tmap) # Thematic maps are geographical maps in which spatial data distributions are visualized

```

#load the copepod data into R studio
```{r}
library(readr)
data <- read_csv("data-for-course/copepods_raw.csv")
data
```
# Plotting the coordinates for the sample
```{r}
library(ggplot2)
ggplot(data) + 
  aes(x = longitude, y = latitude, color = richness_raw) +
  geom_point()
```
# plotting richness
```{r}
ggplot(data, aes(x = latitude, y = richness_raw)) + 
  stat_smooth() + 
  geom_point()
```
# getting it on the maps by putting coordinates
```{r}
library(sf)
sdat <- st_as_sf(data, coords = c("longitude", "latitude"), 
                 crs = 4326)
```

```{r}
?st_as_sf
```

```{r}
crs4326 <- st_crs(4326)
crs4326 # look at the whole CRS
crs4326$Name # pull out just the name of the crs
```
# whole WKT
```{r}
crs4326$wkt # crs in well-known text format
```
# feature collection points
```{r}
sdat
```
# Cartography
```{r}
plot(sdat) # if only plot(sdat) is done it will create a panel for every dataframe.
```

```{r}
plot(sdat["richness_raw"]) # but in sf we use square brackets to select a single variable
```
# Thematic maps for communication
#using tmap
```{r}
tm_shape(sdat) + 
  tm_dots(col = "richness_raw")
```
# to save a file
```{r}
tmap_save(filename = "Richness-map.png", 
          width = 600, height = 600)
```
# loading shapefiles
```{r}
aus <- st_read("data-for-course/spatial-data/Aussie/Aussie.shp")
```

```{r}
shelf <- st_read("data-for-course/spatial-data/aus_shelf/aus_shelf.shp")
```
# checking the data
```{r}
aus
```
# checking the data
```{r}
shelf
```
# Mapping the polygons
```{r}
tm_shape(shelf) + 
  tm_polygons()
```
# combining the data and the map
```{r}
tm_shape(shelf, bbox = sdat) + 
  tm_polygons() +
  tm_shape(aus) + 
  tm_polygons() + 
  tm_shape(sdat) + 
  tm_dots()
```
# Exploring t_map
```{r}
?tmap
```

```{r}
vignette('tmap-getstarted')
```
# do the exercises in tmap
```{r}
data ("World")
tm_shape(World) +
    tm_polygons("HPI")
```
# interactive maps
```{r}
tmap_mode("view")

tm_shape(World) +
    tm_polygons("HPI")
```
#Multiple shapes and layers
```{r}
data(World, metro, rivers, land)

tmap_mode("plot")
## tmap mode set to plotting
tm_shape(land) +
    tm_raster("elevation", palette = terrain.colors(10)) +
tm_shape(World) +
    tm_borders("white", lwd = .5) +
    tm_text("iso_a3", size = "AREA") +
tm_shape(metro) +
    tm_symbols(col = "red", size = "pop2020", scale = .5) +
tm_legend(show = FALSE)
## old-style crs object detected; please recreate object with a recent sf::st_crs()
## old-style crs object detected; please recreate object with a recent sf::st_crs()
## old-style crs object detected; please recreate object with a recent sf::st_crs()
```
```{r}
tmap_save(filename = "world-map.png", 
          width = 1200, height = 1200)
```
# Facets- it can created in 3 ways:
# 1. By assigning multiple variable names to one aesthetic (in this example the first argument of tm_polygons:
```{r}
tmap_mode("view")
tm_shape(World) +
    tm_polygons(c("HPI", "economy")) +
    tm_facets(sync = TRUE, ncol = 2)
```
# 2. By splitting the spatial data with the by argument of tm_facets:
```{r}
tmap_mode("plot")
## tmap mode set to plotting

data(NLD_muni)

NLD_muni$perc_men <- NLD_muni$pop_men / NLD_muni$population * 100

tm_shape(NLD_muni) +
    tm_polygons("perc_men", palette = "RdYlBu") +
    tm_facets(by = "province")
```
# 3. By using the tmap_arrange function:
```{r}
tmap_mode("plot")
## tmap mode set to plotting

data(NLD_muni)
tm1 <- tm_shape(NLD_muni) + tm_polygons("population", convert2density = TRUE)
tm2 <- tm_shape(NLD_muni) + tm_bubbles(size = "population")

tmap_arrange(tm1, tm2)

```
# Basemaps and overlay tile maps
```{r}
tmap_mode("view")
tm_basemap("Stamen.Watercolor") +
tm_shape(metro) + tm_bubbles(size = "pop2020", col = "red") +
tm_tiles("Stamen.TonerLabels")
```
# options and style
```{r}
tmap_mode("plot")
## tmap mode set to plotting

tm_shape(World) +
    tm_polygons("HPI") +
tm_layout(bg.color = "skyblue", inner.margins = c(0, .02, .02, .02))

```

```{r}
tmap_options(bg.color = "lightgreen", legend.text.color = "white")

tm_shape(World) +
    tm_polygons("HPI", legend.title = "Happy Planet Index")
```
# Changing style
```{r}
tmap_style("classic")
## tmap style set to "classic"
## other available styles are: "white", "gray", "natural", "cobalt", "col_blind", "albatross", "beaver", "bw", "watercolor"

tm_shape(World) +
    tm_polygons("HPI", legend.title = "Happy Planet Index")
```
# see what options have been changed
```{r}
tmap_options_diff()
```
# reset the otions back to the default values
```{r}
tmap_options_reset()
```
# Exporting maps
```{r}
tm <- tm_shape(World) +
    tm_polygons("HPI", legend.title = "Happy Planet Index")

## save an image ("plot" mode)
tmap_save(tm, filename = "world_map.png")

## save as stand-alone HTML file ("view" mode)
tmap_save(tm, filename = "world_map.html")
```
# Quick thematic map
```{r}
qtm(World, fill = "HPI", fill.palette = "RdYlGn")
```

```{r}
tmap_tip()
```
# how to import raster data
```{r}
rsst <- rast('data-for-course/spatial-data/MeanAVHRRSST/MeanAVHRRSST.grd')
plot(rsst) # rast will read the raster data in as a SpatRaster object.
```

```{r}
tm_shape(rsst) + 
  tm_raster(palette = "-RdBu", title = "SST")
```


